Towards a wide-coverage grammar for Swedish

### 2011-12-17 doc/report/report.tex
report: evaluation, future work

\chapter{Results and discussion}
\label{sec:results}
\section{Evaluation}
The project has resulted in
\begin{itemize}
\item one large-scale GF lexicon and a program to redo the importation when needed
\item a comparison between GF and other annotations
\item a deeper testing of the Swedish resource grammar and an estimation
of how well GF can be used to describe larger parts of a language
\item a study of how dependent types can be used in the resource grammars
\end{itemize}

The increase of the grammatical coverage is hard to compare, since
The grammar in it self is not at all robust, and one unknown word will cause a whole sentence to 
fail. 
An estimation can be given by looking at the results from running the test suite.
The first grammar could parse almost halv of the sentences, and the final
grammar could parse 65\%. 
When parsing text, we get low results, for this reason. Not yet interesting to talk about
coverage, but focus on quality and ability to scale up. Our goal standarn is so far Talbanken,
but considering that the low coverage of the grammar without statisticts, a comparison would not be
describing anything, to little indata.
Instead we have the luck to have an expert and her presence has increased the standard.
No syntactic theory will ever cover all language, and this should just be seen as an important
fragment.

From the renewed import of Saldo, we got a much better coverage especially considering
that the old version was missing many of the most common words. Bo, göra, se.
However, the lexicon is now to big to use as a normal GF lexicon, since it will request
too much memory to parse most sentences. We will need to handle this in a clever way,
see future work.
%Results from mapping, can probably be improved, bit by bit, depends much on words.
%Valencys would therefore help.  
%Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
%results. Mostly compound nouns, which we can't expect in the dictionary.
%Is too slow, can't be used for parsing right away
%
%Grammar - hard to evaluate automatically, but Elibet is an expert who has been involved
%in the process to verify the solutions. Testing trees against talbanken.
%The big lexicon makes it very slow, eats all memory
%
%Discussion of the results and methods, why are the results good, why are they not better.
%What other methods could have been used? What did we/I expect and what happend?

\section{Future Work}
\label{sec:future}
A part of the project is finished, but the project is not complete, and will never be.
Many interesting future directions. We will first look at the ones which are 
direct consequences of what we did so far. 
\subsection{Enhancments and completitions}
\subsubsection{Grammar}
The are still many constructions that are not covered by the grammar.
As example we can mention
\textbf{object shift}
\enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
a. & jag & ser&  honom&  inte&~~ & b.&   * jag &ser &huset     &inte\\
   & I   & see&  him  &  not & & b.&    I   &see &the house &not\\
\end{tabular}}
\textbf{bare indefinites}
\enumsentence{\shortex{4}
{Vi & ska & köpa & hund} 
{we& will & buy & dog}
{``We will buy a dog"}}
\textbf{måttsattribut}
\enumsentence{
{Antalet bilar, ett antal bilar, ett stort antal}}
Genitive; Not like 'sonen till mig' -> antalet bilars??\\

@@

\section{Conclusion}
>> why is this significant, other grammars can use this Swedish grammar.
>> well known (comp. ling) trade-off between depth and coverage. for now depth, robustness in in future
>> work leads to coverage.
>> reusablilty, freely available.
>> have experiments of how to extract info from talbanken.
%There is obviously a lot do still, will never get finish with NPL.
%Grammar may need even bigger changes, to allow enough but not too much.
%Different from writing a grammar for generation.
%Results from mapping shows that the translation is doable, but becomes harder
%since talbanken don't have formal rules.
%The saldo shows how easy to make use of utomstående resources, promising if
%we want to extract other information. 
%Hard, but not finished!

>>>>

\chapter{Results and discussion}
\label{sec:results}
\section{Evaluation}
The project has resulted in
\begin{itemize}
\item a large-scale GF lexicon and a program to redo the importation when needed
\item a comparison between GF and other annotations
\item a deeper testing of the Swedish resource grammar and an estimation
of how well GF can be used to describe larger parts of a language
\item a study of how dependent types can be used in the resource grammars
\end{itemize}

The grammar has been extended and enhanced, and its current status is
a specialized extension of the resource grammar. When it comes to parsing,
however, we do not get far without robustness.
The grammar in itself is by no means robust, and just one unknown word,
unexpected punctuation
or missing phrase will cause the parsing of the whole sentence to fail. 
Our goal is to parse Talbanken, but considering that no syntactic theory is capable
of covering a whole natural language, our grammar should just be seen as an important
fragment.
A comparison would not be of much value of this state since the coverage of Talbanken
is too small to do be able to get any interesting analysis.\\

An estimation of the improvement can be given by looking at the results from
running the testsuite used for the grammar development.
The sentences given in the testsuite are short and the words are in most cases known
to the lexicon, but there are also constructions that have not been implemented 
during this project.
The first grammar could parse almost half of the sentences, the result for the final
grammar was 65\%. 
Parsing bigger parts of Talbanken would give very low results, for the reasons stated.
% try NPs?
It is thus not yet interesting to talk about coverage, but about the quality and
the ability to scale up, which has so far proved to be good. The constructions
that we have focused on have all been possible to implement, although the amount
of work to get it ~~right has varied.
I further believe that the presence of a Swedish expert has
increased the standard substantially.\\

By the renewed import of Saldo, we doubled the size of the lexicon and thereby
added many of the commonly used irregular words that were missing from the older
version. This is of course a big improvement.
However, the lexical part still requires much work before it can be made use of.
We need valency information to make a good analysis. The lexicon is also too
big to use in a trivial manner, its size makes the incremental parsing algorithm
use too much memory. To solve this, we need to use the lexicon data more cleverly.

%Results from mapping, can probably be improved, bit by bit, depends much on words.
%Valencys would therefore help.  
%Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
%results. Mostly compound nouns, which we can't expect in the dictionary.
%Is too slow, can't be used for parsing right away
%
%Grammar - hard to evaluate automatically, but Elibet is an expert who has been involved
%in the process to verify the solutions. Testing trees against talbanken.
%The big lexicon makes it very slow, eats all memory
%
%Discussion of the results and methods, why are the results good, why are they not better.
%What other methods could have been used? What did we/I expect and what happend?

\section{Future Work}
\label{sec:future}
Although this one part of the project is finished, there are many interesting future
directions. As a natural language processing-project ~ it will hardely veer be entirely
complete.
The future work described in this section is divided into two categories:
the ones aiming at making the parser robust and the ones that can be
seen  as extensions of the work done so far.

\subsection{Enhancments and completitions}
\subsubsection{Grammar}
The are still many constructions that are not covered by the grammar.
As example we can mention
\textbf{object shift}
\enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
a. & jag & ser&  honom&  inte&~~ & b.&   * jag &ser &huset     &inte\\
   & I   & see&  him  &  not & & b.&    I   &see &the house &not\\
\end{tabular}}
\textbf{bare indefinites}
\enumsentence{\shortex{4}
{Vi & ska & köpa & hund} 
{we& will & buy & dog}
{``We will buy a dog"}}
\textbf{måttsattribut}
\enumsentence{
{Antalet bilar, ett antal bilar, ett stort antal}}
Genitive; Not like 'sonen till mig' -> antalet bilars??\\

@@

\section{Conclusion}
>> why is this significant, other grammars can use this Swedish grammar.
>> well known (comp. ling) trade-off between depth and coverage. for now depth,
>> robustness in in future work leads to coverage.
>> reusablilty, freely available.
>> have experiments of how to extract info from talbanken.
By combining GF with a large lexicon, we acquire a deep parser for 
Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
We have developed tools for extending the lexicon with words from Talbanken,
and we show how to make use of the information in the manually tagged
treebank.
The usage of GF allows us to start from
a well-defined system for describing grammar, as well as tools for
parsing.

We are building a parser which is based on grammatical rules rather than
statistics. The parser will therefore validate that
a given sentence is grammatically correct according to the rules defined.
Chosen parts of the parser or grammar may be used in other applications
which deal with controlled natural languages. Automatic language
generation from the grammar is also provided by GF.

%There is obviously a lot do still, will never get finish with NPL.
%Grammar may need even bigger changes, to allow enough but not too much.
%Different from writing a grammar for generation.
%Results from mapping shows that the translation is doable, but becomes harder
%since talbanken don't have formal rules.
%The saldo shows how easy to make use of utomstående resources, promising if
%we want to extract other information. 
%Hard, but not finished!
